Parallel forecasting

### 2020-12-04 r_notebooks/parallel_forecasting.Rmd
Adding files from local work

```{r, setup}
# Packages ####
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(furrr)
library(knitr)

# Options ####
knitr::opts_chunk$set(cache = TRUE)
options(mc.cores = parallel::detectCores() - 1)
rstan_options(auto_write = TRUE)

@@

holidays <- 
  bind_rows(tibble(holiday = "christmas",
                   ds = as.Date(c("2016-12-25", "2017-12-25", "2018-12-25",
                                  "2019-12-25", "2020-12-25", "2021-12-25")),
                   lower_window = -1, 
                   upper_window = 1),

            tibble(holiday = "easter", # Sunday
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -2, 
                   upper_window = 1))

countRegional <- function(data, region, datename = datename, unit = Quantity){
 # Function to create tsibble object for a count of `unit` per `datename`
  # e.g. countRegional(oc, Locality, datename, Quantity) is the `Quantity`
  # per `datename` grouped by `Locality` from `oc`. `Quantity` is corrected for
  # the number of days in each month.
  region <- enquo(region)
  datename <- enquo(datename)
  unit <- enquo(unit)

  data %>% 
  mutate(regional_unit = as_factor(set_names(!!region))) %>% 
  group_by(regional_unit, !!datename) %>% 
  summarise(n = sum(!!unit)) %>% 
  mutate(n = case_when(str_detect(!!datename, pattern = "-01-") ~ n/31,
                       str_detect(!!datename, pattern = "-02-") ~ n/28.25,
                       str_detect(!!datename, pattern = "-03-") ~ n/31,
                       str_detect(!!datename, pattern = "-04-") ~ n/30,
                       str_detect(!!datename, pattern = "-05-") ~ n/31,
                       str_detect(!!datename, pattern = "-06-") ~ n/30,
                       str_detect(!!datename, pattern = "-07-") ~ n/31,
                       str_detect(!!datename, pattern = "-08-") ~ n/31,
                       str_detect(!!datename, pattern = "-09-") ~ n/30,
                       str_detect(!!datename, pattern = "-10-") ~ n/31,
                       str_detect(!!datename, pattern = "-11-") ~ n/30,
                       str_detect(!!datename, pattern = "-12-") ~ n/31)) %>% 
  ungroup() %>% 
  as_tsibble(index = !!datename, key = regional_unit)
}

regionalPlot <- function(region, data){

  ## Prophet forecast - can this be taken outside of the function?
      ## no, when it is taken outside of the model, the forecast is becomes 
      ## much worse 2020-11-14
  df <- data %>%
    filter(regional_unit == region & datename < ymd("2020-03-23")) %>%
    transmute(ds = datename, y = n)

  m <- prophet::prophet(df, holidays = holidays,
                        seasonality.mode = "additive",# mcmc.samples = 1000,
                        weekly.seasonality = FALSE, daily.seasonality = FALSE)
  future <- make_future_dataframe(m, periods = 5, freq = 'month')
  fcst <- predict(m, future)

  ## Plotting
  plot <- fcsts %>% 
    filter(regional_unit == region) %>% 
    fabletools::autoplot(data, level = 80) + # fable forecasts over observed data
    geom_line(aes(y = c(rep(NA,50), fcst[51:55,]$yhat)), # prophet prediction
              col = "blue", lty = 3) +
    geom_ribbon(aes(ymin = c(rep(NA,50), fcst[51:55,]$yhat_lower), # prophet confidence
                    ymax = c(rep(NA,50), fcst[51:55,]$yhat_upper)),
                alpha = .3, fill = "blue") +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") + # add line to mark the start of lockdown
    labs(title = region)

  ## Results
  result <- list("plots" = plot)#, "prophet" = fcst)
  return(result)
}

fableModels <- function(data){
  data %>% 
    model(trend_model1 = TSLM(n ~ trend()),#time series linear model
          trend_model2 = TSLM(n ~ trend() + season("year")),
          ets1 = ETS(n ~ trend()),
          ets2 = ETS(n ~ trend() + season("A")),
          ets3 = ETS(n ~ trend() + season("M")),
          arima = ARIMA(n, stepwise = FALSE, approximation = FALSE),
          comb1 = combination_model(TSLM(n ~ trend()), 
                                    ETS(n ~ trend())),
          comb2 = combination_model(TSLM(n ~ trend() + season("year")), 
                                    ETS(n ~ trend() + season("A"))),
          stl_dcmp1 = decomposition_model(STL(n ~ trend(),
                                              iterations = 1000, 
                                              robust = FALSE),
                                          SNAIVE(season_adjust)),
          stl_dcmp2 = decomposition_model(STL(n ~ trend() + season(),
                                              iterations = 1000, 
                                              robust = FALSE),
                                          SNAIVE(season_adjust)),
          s_naive = SNAIVE(n))
}

regionalFits <- function(data, region){
  # Return various model objects for each region
  region <- enquo(region)

  data %>% 
    filter(regional_unit == !!region) %>% 
    mutate(datename = yearmonth(datename)) %>% 
    fableModels()
}

prophetModels <- function(data, region){
  # Return various model objects for each region
  region <- enquo(region)

  oc_prophet_model1 <- 
    fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model2 <- 
    fable.prophet::prophet(n ~ season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model3 <- 
    fable.prophet::prophet(n ~ growth("linear",) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model4 <- 
    fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model5 <- 
    fable.prophet::prophet(n ~ growth("linear",) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model6 <- 
    fable.prophet::prophet(n ~ season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model7 <- 
    fable.prophet::prophet(n ~ holiday(holidays))
  oc_prophet_model8 <- 
    fable.prophet::prophet(n)

  data %>% 
    as_tsibble(index = datename) %>% 
    model(oc_prophet_model1, oc_prophet_model2, oc_prophet_model3, 
          oc_prophet_model4, oc_prophet_model5, oc_prophet_model6, 
          oc_prophet_model7, oc_prophet_model8, .safely = TRUE)
}

regionalFitsProphet <- function(data, region){
  # Return various model objects for each region
  region <- enquo(region)

  oc_prophet_model1 <- 
    fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model2 <- 
    fable.prophet::prophet(n ~ season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model3 <- 
    fable.prophet::prophet(n ~ growth("linear",) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model3 <- 
    fable.prophet::prophet(n ~ growth("linear", n_changepoints = 0) +
                             season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model5 <- 
    fable.prophet::prophet(n ~ growth("linear",) +
                             season("year", type = "additive") + holiday(holidays))
  oc_prophet_model6 <- 
    fable.prophet::prophet(n ~ season("year", type = "multiplicative") + holiday(holidays))
  oc_prophet_model7 <- 
    fable.prophet::prophet(n ~ holiday(holidays))
  oc_prophet_model8 <- 
    fable.prophet::prophet(n)

  data %>% 
    filter(regional_unit == !!region) %>% 
    as_tsibble(index = datename) %>% 
    model(oc_prophet_model1, oc_prophet_model2, oc_prophet_model3, 
          oc_prophet_model4, oc_prophet_model5, oc_prophet_model6, 
          oc_prophet_model7, oc_prophet_model8, .safely = TRUE)
}

@@

## Smaller GP data set to develop process
regions <- unique(as.character(gp$PracticeID))
dev_data <- 
  oc_monthly_quantity_gp_pre_cv19 %>% 
  filter(regional_unit %in% sample(regions, 50))

@@

@@

```{r, eval=FALSE}
system.time(fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFits(dev_data, .x)))

@@

system.time(
fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFits(dev_data, .x), 
             .options = furrr_options(seed = TRUE)))

@@

```{r}
system.time(proph_fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFitsProphet(dev_data, .x)))

@@

system.time(proph_fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFitsProphet(dev_data, .x)))

@@

```{r, eval=FALSE}
plan(multisession)
system.time(
  proph_fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
                               .f = ~regionalFitsProphet(dev_data, .x), 
                               .options = furrr_options(seed = TRUE)))

@@

```{r}
regionalFitsJoint <- function(data){
  regions <- unique(data$regional_unit)

  plan(multisession)
  fable_fit <- 
    future_map_dfr(.x = regions, 
                   .f = ~regionalFits(data, .x), 
                   .options = furrr_options(seed = TRUE)) 
  message("fable fits")

  plan(sequential)

  prophet_fit <- 
    map_dfr(.x = regions, 
            .f = ~regionalFitsProphet(data, .x))
  message("prophet fits")

  message("prophet forecasts")

  joint_fit <- 
    left_join(fable_fit, prophet_fit, by = "regional_unit")

  return(joint_fit)
}
```

@@

plan(multisession)
fable_fit <-
  future_map_dfr(.x = regions,
                  .f = ~regionalFits(filter(dev_data_tr, .id == 1), .x),
                  .options = furrr_options(seed = TRUE))

@@

  print(paste("fable regional fit", window))
  future_map_dfr(.x = regions,
                 .f = ~regionalFits(filter(data, .id == window), .x),
                 .options = furrr_options(seed = TRUE))
}

@@

```{r}
tictoc::tic()
regions <- unique(dev_data$regional_unit)
cv_dist <- 6
init <- 36
step <- 1
data_test <- mutate(filter(dev_data, regional_unit == regions[[1]]),
                  datename = yearmonth(datename))
data_tr <- dev_data %>% 
  filter(regional_unit == regions[1]) %>% 
  select(regional_unit, datename, n) %>% 
  mutate(datename = yearmonth(datename)) %>% 
  fill_gaps() %>% 
  slice(1:(n()-cv_dist), .preserve = TRUE) %>% 
  stretch_tsibble(.init = init, .step = step)

bind_rows(data_tr %>% 
  fableModels() %>% 
  forecast(h = cv_dist) %>% 
  accuracy(data_test),
data_tr %>% 
  prophetModels() %>% 
  forecast(h = cv_dist) %>% 
  accuracy(data_test)) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1)

tictoc::toc()

regionalCV <- function(data, region, cv_dist = 6, init = 36, step = 1){
  # Data must be filtered outside of the function
  data <-
    filter(data, regional_unit == region)
  data_test <- 
    data %>% 
    mutate(datename = yearmonth(datename))

  data_tr <- data %>% 
    select(regional_unit, datename, n) %>% 
    mutate(datename = yearmonth(datename)) %>% 
    fill_gaps() %>% 
    slice(1:(n()-cv_dist), .preserve = TRUE) %>% 
    stretch_tsibble(.init = init, .step = step)

  plan(multisession)
  fable_fits <- 
    data_tr %>% 
    fableModels() %>% 
    forecast(h = cv_dist) %>% 
    accuracy(data_test)
  plan(sequential)

  prophet_fits <- 
    data_tr %>% 
    prophetModels() %>% 
    forecast(h = cv_dist) %>% 
    accuracy(data_test)

  bind_rows(fable_fits, prophet_fits) %>% 
    arrange(RMSE, MAE) %>% 
    slice(1)
}

regions <- unique(dev_data$regional_unit)
# plan(multisession)
system.time(best_models <- map_dfr(.x = unique(dev_data$regional_unit),
        .f = ~regionalCV(data = dev_data, region = .x),
        .id = "a"))
# plan(sequential)

best_models %>% 
  group_by(.model) %>% 
  count %>% 
  arrange(desc(n))

best_models %>% 
  ggplot(aes(x = MASE)) +
  geom_histogram()
```

* 1 regional_unit
  * Mean time `r mean(c(78.5, 77.7))` s w/o prophet
  * Mean time `r mean(c(142.466, 147.631))` s w prophet
* 2 regional units
  * min = 671.0648, lq = 691.4213, mean = 699.9314, median = 700.2545, uq = 710.868, max = 726.0485,

@@

bestModelFit(accuracies, rmse)

@@

regionalFitsJoint <- function(data){
  regions <- unique(data$regional_unit)

  plan(multisession)
  fable_fit <- 
    future_map_dfr(.x = regions, 
                   .f = ~regionalFits(data, .x), 
                   .options = furrr_options(seed = TRUE)) 
  message("fable fits")

  fable_fcst <- forecast(fable_fit, h = "5 months") %>% 
    as_tibble()

  message("fabel forecasts")

  plan(sequential)

  prophet_fit <- 
    map_dfr(.x = regions, 
            .f = ~regionalFitsProphet(data, .x))
  message("prophet fits")

  prophet_fcst <- forecast(prophet_fit, h = "5 months") %>% 
    as_tibble()
  message("prophet forecasts")

  joint_fit <- 
    left_join(fable_fit, prophet_fit, by = "regional_unit")
  joint_fcst <- 
    bind_rows(fable_fcst, prophet_fcst) %>% 
    mutate(lb_95 = quantile(n, .025),
           lb_80 = quantile(n, .1),
           med = median(n),
           ub_80 = quantile(n, .9),
           ub_95 = quantile(n, .975),
           n = NULL)

  return(list(fits = joint_fit, forecasts = joint_fcst))
}

>>>>

```{r, setup}
# Packages ####
library(tidyverse)
library(lubridate)
library(tsibble)
library(rstan) # needed to load prophet
library(prophet)
library(fable)
library(fable.prophet)
library(feasts)
library(fasster)
library(furrr)
library(knitr)

# Options ####
knitr::opts_chunk$set(cache = TRUE)
options(mc.cores = parallel::detectCores() - 1)
rstan_options(auto_write = TRUE)
options(future.rng.onMisue = "ignore")

@@

source("../r_scripts/functions.r")

@@

## Smaller GP data set to develop process
regions <- unique(as.character(gp$PracticeID))
dev_data <- 
  oc_monthly_quantity_gp_pre_cv19 %>% 
  filter(regional_unit %in% sample(regions, 5))

@@

```{r}
dev_data %>% 
  fableModels()
```

@@

```{r, eval=FALSE}
system.time(fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFableFits(dev_data, .x)))

@@

system.time(
fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalFableFits(dev_data, .x), 
             .options = furrr_options(seed = TRUE)))

@@

```{r}
system.time(proph_fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalProphetFits(dev_data, .x)))

@@

system.time(proph_fit2 <- map_dfr(.x = unique(dev_data$regional_unit), 
             .f = ~regionalProphetFits(dev_data, .x)))

@@

```{r, eval=FALSE}
plan(multisession)
system.time(
  proph_fit3 <- future_map_dfr(.x = unique(dev_data$regional_unit), 
                               .f = ~regionalProphetFits(dev_data, .x), 
                               .options = furrr_options(seed = TRUE)))

@@

```{r}

@@

plan(multisession)
fable_fit <-
  future_map_dfr(.x = regions,
                  .f = ~regionalFableFits(filter(dev_data_tr, .id == 1), .x),
                  .options = furrr_options(seed = TRUE))

@@

  print(paste("fable regional fit", window))
  future_map_dfr(.x = regions,
                 .f = ~regionalFableFits(filter(data, .id == window), .x),
                 .options = furrr_options(seed = TRUE))
}

@@

```{r}

fableModels(dev_data)
regionalFableFits(dev_data, Arfon)

system.time(best_models_1 <- regionalCV(dev_data))
system.time(best_models_2 <- regionalCV(dev_data, init = 30))
system.time(best_models_3 <- regionalCV(dev_data, step = 2))

system.time(best_oc_models1 <- regionalCV(oc_monthly_quantity_gp_pre_cv19))

best_oc_models1 %>% 
  saveRDS(file = "../data/oc_gp_cv_models.rds")

best_models_1 %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  ggplot(aes(x = RMSE)) +
  geom_histogram()

best_models_1 %>% 
  group_by(regional_unit) %>% 
  arrange(RMSE, MAE) %>% 
  slice(1, .preserve = TRUE) %>% 
  group_by(.model) %>% 
  count() %>% 
  arrange(desc(n))
```

* 1 regional_unit
  * Mean time `r mean(c(78.5, 77.7))` s w/o prophet
  * Mean time `r mean(c(142.466, 147.631))` s w prophet
* 2 regional units
  * min = 671.0648, lq = 691.4213, mean = 699.9314, median = 700.2545, uq = 710.868, max = 726.0485, 
* 100 regional units
  * Mean time `r mean(c(66, 61))` mins
* 400 regional units
  * `r lubridate::seconds_to_period(29663.27)` hours

### prophet::cross_validation()

```{r}
cross_validation(select(oc_gp_cluster_fcts$fits, c(regional_unit, starts_with("oc_proph"))), horizon = .5, units = "years", initial = 3)
```

@@

bestModelFit(accuracies, rmse)
```

@@
